---
title: "Simulacrum Logistic Regression"
author: "Matt Grant"
date: "2023-03-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
options(scipen=999)

```

```{r}
simulacrum <- readRDS(file = "simulacrum_av_data_clean.rds")

simulacrum <- simulacrum %>%
  select(!c(CancerSiteLevel3Code.y, CancerSiteLevel3CodeDesc.y, CancerSiteLevel3Desc.y, CancerSiteLevel3PublishDesc.y, SEX.y, AGE_GROUP,CancerSiteLevel3Code.x, CancerSiteLevel3Desc.x, CancerSiteLevel3PublishDesc.x, CancerSiteLevel3CodeDesc.x, ICD10_INITIAL, LINKNUMBER.y))
```

```{r}
t_simulacrum <- simulacrum
```

# Cleaning

## Initial cleaning

The dataset I'm using is a cleaned version of the simulacrum datasets (`sim_av_patient` and `sim_av_tumour`). It needs to be cleaned up a bit more before running a logistic regression.

Steps to take:

-   Remove any patient with multiple tumours

-   Remove all but breast, bowel, and lung cancer patients

-   Remove any columns that are dates

-   Remove any columns that are related to death ('causeofdeath') -- except the newvitalstatus, which is your target

-   Remove patients where `newvitalstatus` is 'X'

-   Change empty values to NA

-   Remove `ER_STATUS` and `PR_STATUS` (keeping `ER_SCORE` and `PR_SCORE` for prediction)

-   Remove GLEASON measures

```{r}
t_simulacrum <- t_simulacrum %>%
  distinct(PATIENTID, .keep_all = TRUE) %>%
  filter(
    #Breast cancer
    SITE_ICD10_O2_3CHAR == "C50" | 
    #Bowel cancer
    SITE_ICD10_O2_3CHAR == "C18" |
    SITE_ICD10_O2_3CHAR == "C19" |
    SITE_ICD10_O2_3CHAR == "C20" |
    SITE_ICD10_O2_3CHAR == "C21" | 
    #Lung cancer
    SITE_ICD10_O2_3CHAR == "C33" |
    SITE_ICD10_O2_3CHAR == "C34") %>% 
  select(!c(VITALSTATUSDATE, VITALSTATUSDATE_YEAR, DIAGNOSISDATEBEST, DIAGNOSIS_YEAR, DATE_FIRST_SURGERY, DEATHCAUSECODE_1A, DEATHCAUSECODE_1B, DEATHCAUSECODE_1C, DEATHCAUSECODE_2, DEATHCAUSECODE_UNDERLYING, DEATHLOCATIONCODE, LINKNUMBER.x, SITE_ICD10_O2, MORPH_ICD10_O2, BEHAVIOUR_ICD10_O2, STAGE_BEST_SYSTEM,  GLEASON_PRIMARY, GLEASON_SECONDARY, GLEASON_TERTIARY, GLEASON_COMBINED, SCREENINGSTATUSFULL_CODE, CNS, CREG_CODE)) %>%
  filter(NEWVITALSTATUS != 'X') 
t_simulacrum <- na_if(t_simulacrum, '')
t_simulacrum <- t_simulacrum %>%
  select(!PATIENTID)

table(t_simulacrum$SITE_ICD10_O2_3CHAR)
```

## Cleaning NA values

```{r}
t_simulacrum %>% summarise(across(everything(), ~ sum(is.na(.))))

```

### STAGE_BEST

```{r}
table(t_simulacrum$STAGE_BEST, useNA = 'ifany')
```

| Value | Description       |
|-------|-------------------|
| 0     | 0                 |
| 1     | 1                 |
| 2     | 2                 |
| 3     | 3                 |
| 4     | 4                 |
| 5     | Outdate/invalid   |
| 6     | Not stageable     |
| ?     | Insufficient Info |
| U     | Unstageable       |

Group 1A, 1A1, etc. into their initial integer to align with dictionary

Changing `NA` and `6` to `?`

Setting `0` as the factor level reference.

```{r}
t_simulacrum <- t_simulacrum %>%
  mutate(STAGE_BEST = substring(t_simulacrum$STAGE_BEST, 1, 1))

t_simulacrum$STAGE_BEST <- replace_na(t_simulacrum$STAGE_BEST, '?')
t_simulacrum$STAGE_BEST <- t_simulacrum$STAGE_BEST %>%
  replace(t_simulacrum$STAGE_BEST == '6', '?') %>%
  replace(t_simulacrum$STAGE_BEST == 'X', '?')
t_simulacrum$STAGE_BEST <- factor(t_simulacrum$STAGE_BEST, levels = c('0','1', '2', '3', '4', 'U', '?'))

table(t_simulacrum$STAGE_BEST, useNA = 'always')


STAGE_BEST_TABLE <- round(prop.table(table(t_simulacrum$NEWVITALSTATUS,
                                           t_simulacrum$STAGE_BEST), 
                                     margin = 2), 
                          2)
STAGE_BEST_TABLE
round(prop.table(table(t_simulacrum$STAGE_BEST)), 2)
```

### T_BEST

```{r}
table(t_simulacrum$T_BEST, useNA = "ifany")
```

Recode NA to 0 which is "Main tumour cannot be found". The reference grade for this feature will be `1.`

```{r}
t_simulacrum$T_BEST <- replace_na(t_simulacrum$T_BEST, '0')
t_simulacrum$T_BEST <- factor(t_simulacrum$T_BEST, levels = c('1', '2', '3', '4', '0'))

table(t_simulacrum$T_BEST, useNA = 'always')
```

### M_BEST

```{r}
table(t_simulacrum$M_BEST, useNA = 'always')
```

Measure explanation:

Distant metastasis (M)

-   MX: Metastasis cannot be measured.

-   M0: Cancer has not spread to other parts of the body.

-   M1: Cancer has spread to other parts of the body.

So recode NA in to X.

Reference factor for this feature is 0, cancer has not spread to other parts of the body

```{r}
t_simulacrum$M_BEST <- replace_na(t_simulacrum$M_BEST, 'X')
t_simulacrum$M_BEST <- factor(t_simulacrum$M_BEST, levels = c('0', '1', 'X'))
table(t_simulacrum$M_BEST, useNA = 'always')
```

### N_BEST

```{r}
table(t_simulacrum$N_BEST, useNA = "always")
```

Regional lymph nodes (N)

-   NX: Cancer in nearby lymph nodes cannot be measured.

-   N0: There is no cancer in nearby lymph nodes.

-   N1, N2, N3: Refers to the number and location of lymph nodes that contain cancer. The higher the number after the N, the more lymph nodes that contain cancer.

Reference factor for `N_BEST` will be 0 - there is no cancer in nearby lymph nodes.

```{r}
#Recode NA to X
t_simulacrum$N_BEST <- replace_na(t_simulacrum$N_BEST, 'X')
t_simulacrum$N_BEST <- factor(t_simulacrum$N_BEST, levels = c('0', '1', '2', '3',  'X'))
table(t_simulacrum$N_BEST, useNA = 'always')
```

### PERFORMANCESTATUS

```{r}
table(t_simulacrum$PERFORMANCESTATUS, useNA = "always")
```

| **GRADE** |                                                                **ECOG PERFORMANCE STATUS**                                                                |
|:---------:|:---------------------------------------------------------------------------------------------------------------------------------------------------------:|
|     0     |                                      Fully active, able to carry on all pre-disease performance without restriction                                       |
|     1     | Restricted in physically strenuous activity but ambulatory and able to carry out work of a light or sedentary nature, e.g., light house work, office work |
|     2     |              Ambulatory and capable of all selfcare but unable to carry out any work activities; up and about more than 50% of waking hours               |
|     3     |                                 Capable of only limited selfcare; confined to bed or chair more than 50% of waking hours                                  |
|     4     |                                    Completely disabled; cannot carry on any selfcare; totally confined to bed or chair                                    |

We also know that 9 is used as a dump code.

Reference factor for this feature is `0` - fully active

```{r}
#recode m values to whole integers
t_simulacrum$PERFORMANCESTATUS <- t_simulacrum$PERFORMANCESTATUS %>%
  replace(t_simulacrum$PERFORMANCESTATUS == '1m', 1) %>%
  replace(t_simulacrum$PERFORMANCESTATUS == '2m', 2) %>%
  replace(t_simulacrum$PERFORMANCESTATUS == '3m', 3) %>%
  replace(t_simulacrum$PERFORMANCESTATUS == '4m', 4) 
  #recode NA to 5
t_simulacrum$PERFORMANCESTATUS <- replace_na(t_simulacrum$PERFORMANCESTATUS, '5')
t_simulacrum$PERFORMANCESTATUS <- factor(t_simulacrum$PERFORMANCESTATUS, levels = c('0', '1', '2', '3', '4', '5', '9'))
table(t_simulacrum$PERFORMANCESTATUS, useNA = "always")
```

### GRADE

```{r}
table(t_simulacrum$GRADE, useNA = 'always')
```

-   Grade X: Grade cannot be assessed (undetermined grade)

-   Grade 1: Well differentiated (low grade)

-   Grade 2: Moderately differentiated (intermediate grade)

-   Grade 3: Poorly differentiated (high grade)

-   Grade 4: Undifferentiated (high grade)

Since x is "undetermined" we can put all the NA in there. We will use `G1` as the reference factor level for this feature.

```{r}
t_simulacrum$GRADE <- replace_na(t_simulacrum$GRADE, 'GX')
t_simulacrum$GRADE <- factor(t_simulacrum$GRADE, levels = c('G1', 'G2', 'G3', 'G4', 'GX'))
table(t_simulacrum$GRADE, useNA = 'always')
```

### ETHNICITY

```{r}
table(t_simulacrum$ETHNICITY, useNA = 'always')
```

|     |                               |
|-----|-------------------------------|
| 0   | WHITE                         |
| 8   | 8 OTHER                       |
| A   | WHITE BRITISH                 |
| B   | WHITE IRISH                   |
| C   | ANY OTHER WHITE BACKGROUND    |
| D   | WHITE AND BLACK CARIBBEAN     |
| E   | MIXED WHITE AND BLACK AFRICAN |
| F   | MIXED WHITE AND ASIAN         |
| G   | ANY OTHER MIXED BACKGROUND    |
| H   | ASIAN INDIAN                  |
| J   | ASIAN PAKISTANI               |
| K   | ASIAN BANGLADESHI             |
| L   | ANY OTHER ASIAN BACKGROUND    |
| M   | BLACK CARIBBEAN               |
| N   | BLACK AFRICAN                 |
| P   | ANY OTHER BLACK BACKGROUND    |
| R   | CHINESE                       |
| S   | ANY OTHER ETHNIC GROUP        |
| X   | NOT KNOWN                     |
| Z   | NOT STATED                    |

Recoding `NA` as `X` since that is akin to "not known". `WHITE BRITISH` will be the reference factor level for this feature as it is the biggest group.

```{r}
t_simulacrum$ETHNICITY <- replace_na(t_simulacrum$ETHNICITY, 'X')
t_simulacrum$ETHNICITY <- t_simulacrum$ETHNICITY %>%
  replace(t_simulacrum$ETHNICITY == 0, 'C') %>%
  replace(t_simulacrum$ETHNICITY == 8, 'C') 
t_simulacrum$ETHNICITY <- factor(t_simulacrum$ETHNICITY, levels = c('A', 'B', 'C', 'D',  'E', 'F', 'G', 'H', 'I', 'J', 'K', 'L', 'M', 'N', 'P', 'R', 'S', 'X', 'Z'))
table(t_simulacrum$ETHNICITY, useNA = 'always')
```

### ER_STATUS

```{r}
table(t_simulacrum$ER_SCORE, useNA = 'ifany')
table(t_simulacrum$ER_STATUS, useNA = 'ifany')
table(t_simulacrum$ER_STATUS, t_simulacrum$ER_SCORE, useNA = 'always')
```

NHS data dictionary only lists 3 values for `ER_STATUS` (Estrogen Receptor Status): `P` (positive), `N` (negative), `X` (no test performed)

Can comfortable put `Pm` into `P` although it has no associated score, `U` could feasibly be "untested", `B` is listed as "borderline" in `HER2_STATUS` so likely the same here.

We are using `N` - negative as the reference factor level for this feature.

```{r}
t_simulacrum$ER_STATUS <- replace_na(t_simulacrum$ER_STATUS, 'X') %>%
  replace(t_simulacrum$ER_STATUS == 'Pm', 'P') %>%
  replace(t_simulacrum$ER_STATUS == 'U', 'X')
t_simulacrum$ER_STATUS <- factor(t_simulacrum$ER_STATUS, levels = c('N', 'B', 'P', 'X'))
table(t_simulacrum$ER_STATUS, useNA = 'always')
#drop ER_SCORE 
t_simulacrum <- t_simulacrum %>%
  select(!c(ER_SCORE))
```

### PR_STATUS

```{r}
table(t_simulacrum$PR_SCORE, useNA = 'always')
table(t_simulacrum$PR_STATUS, useNA = 'always')
table(t_simulacrum$PR_STATUS, t_simulacrum$PR_SCORE, useNA = 'always')
```

Very similar makeup to `ER_STATUS` and `ER_SCORE` so going to do the same with them.

We are using `N` as the reference factor level for this feature.

```{r}
t_simulacrum$PR_STATUS <- replace_na(t_simulacrum$PR_STATUS, 'X') %>%
  replace(t_simulacrum$PR_STATUS == 'Pm', 'P') %>%
  replace(t_simulacrum$PR_STATUS == 'U', 'X')
t_simulacrum$PR_STATUS <- factor(t_simulacrum$PR_STATUS, levels = c('N', 'B', 'P', 'X'))

table(t_simulacrum$PR_STATUS, useNA = 'always')
#drop ER_SCORE 
t_simulacrum <- t_simulacrum %>%
  select(!c(PR_SCORE))
```

### HER2_STATUS

```{r}
table(t_simulacrum$HER2_STATUS, useNA = 'ifany')
```

Similar to the `ER_STATUS` and `PR_STATUS`.

NHS data dictionary for this code lists `B` as "borderline".

We are using `N` - negative as the reference factor level for this feature.

```{r}
t_simulacrum$HER2_STATUS <- replace_na(t_simulacrum$HER2_STATUS, 'X') %>%
  replace(t_simulacrum$HER2_STATUS == 'Pm', 'P') %>%
  replace(t_simulacrum$HER2_STATUS == 'U', 'X')
t_simulacrum$HER2_STATUS <- factor(t_simulacrum$HER2_STATUS, levels = c('N', 'B', 'P', 'X'))

table(t_simulacrum$HER2_STATUS, useNA = 'always')

```

### CANCERCAREPLANINTENT

```{r}
table(t_simulacrum$CANCERCAREPLANINTENT, useNA = 'ifany')
```

|     |                     |
|-----|---------------------|
| 9   | Not known           |
| C   | Curative            |
| X   | No active treatment |
| Z   | Non Curative        |

Dropping all `NA` into `9` "Not known".

We will use `X` - no active treatment as the reference factor level for this feature.

```{r}
t_simulacrum$CANCERCAREPLANINTENT <- replace_na(t_simulacrum$CANCERCAREPLANINTENT, '9') %>%
  replace(t_simulacrum$CANCERCAREPLANINTENT == 'Cm', 'C') %>%
  replace(t_simulacrum$CANCERCAREPLANINTENT == 'Zm', 'Z')
t_simulacrum$CANCERCAREPLANINTENT <- factor(t_simulacrum$CANCERCAREPLANINTENT, levels = c('X', 'C', 'Z', '9'))

table(t_simulacrum$CANCERCAREPLANINTENT, useNA = 'always')
```

### ACE27

```{r}
table(t_simulacrum$ACE27, useNA = 'ifany')
```

Going to recode `Xm` categories into respective integer, and put `NA` into `9` as it is not listed in the data dictionary.

| Code | Description |
|:----:|:-----------:|
|  0   |    None     |
|  1   |    Mild     |
|  2   |  Moderate   |
|  3   |   Severe    |

The reference factor level for this feature is `0` - none.

```{r}
t_simulacrum$ACE27 <- replace_na(t_simulacrum$ACE27, '9') %>%
  replace(t_simulacrum$ACE27 == '1m', '1') %>%
  replace(t_simulacrum$ACE27 == '2m', '2') %>%
  replace(t_simulacrum$ACE27 == '3m', '3')
t_simulacrum$ACE27 <- factor(t_simulacrum$ACE27, levels = c('0', '1', '2', '3', '9'))

table(t_simulacrum$ACE27, useNA = 'always')
```

### LATERALITY

```{r}
table(t_simulacrum$LATERALITY, useNA = 'ifany') 
```

|     |           |
|-----|-----------|
| L   | Left      |
| R   | Right     |
| M   | Midline   |
| B   | Bilateral |

Going to recode `NA` into `X`.

As `L` is the largest group, it will be used as the reference factor level for this feature.

```{r}
t_simulacrum$LATERALITY <- replace_na(t_simulacrum$LATERALITY, 'X')
t_simulacrum$LATERALITY <- factor(t_simulacrum$LATERALITY, levels = c('L', 'M', 'R', 'B', 'X'))

table(t_simulacrum$LATERALITY, useNA = 'always')
```

### NEWVITALSTATUS

```{r}
t_simulacrum$NEWVITALSTATUS <- as.factor(t_simulacrum$NEWVITALSTATUS)
t_simulacrum$QUINTILE_2015 <- as.factor(t_simulacrum$QUINTILE_2015)

```

`A` - Alive is the reference value for this factor.

## Cancer type

Create a cancer type field that aggregates the relevant cancer types

`C18, C19, C20, C21 - 'Bowel'`

`C33, C34 - 'Lung'`

`C50 - 'Breast'`

Setting `Breast` as the reference category as it is the largest

```{r}
t_simulacrum <- t_simulacrum %>%
  mutate(CANCER_TYPE = case_when(SITE_ICD10_O2_3CHAR == 'C18' ~ 'Bowel',
                                 SITE_ICD10_O2_3CHAR == 'C19' ~ 'Bowel',
                                 SITE_ICD10_O2_3CHAR == 'C20' ~ 'Bowel',
                                 SITE_ICD10_O2_3CHAR == 'C21' ~ 'Bowel',
                                 SITE_ICD10_O2_3CHAR == 'C33' ~ 'Lung',
                                 SITE_ICD10_O2_3CHAR == 'C34' ~ 'Lung',
                                 SITE_ICD10_O2_3CHAR == 'C50' ~ 'Breast'))

table(t_simulacrum$CANCER_TYPE)
t_simulacrum$CANCER_TYPE <- factor(t_simulacrum$CANCER_TYPE, levels = c('Breast', 'Bowel', 'Lung')) 
t_simulacrum <- t_simulacrum %>%
  select(!SITE_ICD10_O2_3CHAR)
```

# Logistical Regression

The aim is to build a logistical regression model that can predict the vital status (`NEWVITALSTATUS`) of breast cancer patients.

## Setting out test data

```{r}
#Build out the sample training and testing sets
set.seed(123)
sample <- sample(c(TRUE, FALSE), nrow(t_simulacrum), replace = T, prob = c(0.7,0.3))
training_data <- t_simulacrum[sample, ]
test_data <- t_simulacrum[!sample, ]
```

Check that the sampling gives similar proportions of alive/dead in the training and test data sets

```{r}
round(prop.table(table(t_simulacrum$NEWVITALSTATUS)) * 100, 1)
round(prop.table(table(training_data$NEWVITALSTATUS)) * 100, 1)
round(prop.table(table(test_data$NEWVITALSTATUS)) * 100, 1)
```

## Naive model

First we need to build the naive model, which has no predictive variables and will act as the baseline for the actual predictive models we build.

```{r}
#create model
naive_model <- glm(NEWVITALSTATUS ~ 1, data = training_data, family=binomial)
summary(naive_model)
```

```{r}
#use model to predict outcomes in training data
naive_model_train_prob <- naive_model %>%
  predict(type = "response")

#Use model to predict outcomes of test data
naive_model_test_prob <- naive_model %>% 
  predict(test_data, type = "response")
naive_model_pred_classes <- ifelse(naive_model_test_prob > 0.5, 
                                                          "D", 
                                                          "A")

#create confusion matrix
naive_model_confusion_matrix <- table(naive_model_pred_classes, test_data$NEWVITALSTATUS)
print(paste0("Confusion matrix"))
naive_model_confusion_matrix
TN = naive_model_confusion_matrix[1, 1]
FN = 0
FP = naive_model_confusion_matrix[1, 2]
TP = 0
paste("TN:", TN, "FN:", FN, "FP:", FP, "TP:", TP)
Accuracy = (TP + TN)/(TP + FP + TN + FN)
Precision = TP/(TP + FP)
Recall = TP/(TP + FN)
F1_score = 2 * ((Precision * Recall) / (Precision + Recall))

print(paste("Test Accuracy:", Accuracy))
print(paste("Precision:", round(Precision, 3)))
print(paste("Recall:", round(Recall, 3)))
print(paste("F1_score:", round(F1_score, 3)))

Naive_Accuracy = Accuracy
```

## Model 1 - CANCER_TYPE

Using `CANCER_TYPE`.

```{r}
#create model
model_1 <- glm(NEWVITALSTATUS ~ CANCER_TYPE, data = training_data, family=binomial)
summary(model_1)
```

```{r}
#use model to predict outcomes in training data
model_1_train_prob <- model_1 %>%
  predict(type = "response")

#Use model to predict outcomes of test data
model_1_test_prob <- model_1 %>% 
  predict(test_data, type = "response")
model_1_pred_classes <- ifelse(model_1_test_prob > 0.5, 
                                                          "D", 
                                                          "A")

#create confusion matrix
model_1_confusion_matrix <- table(model_1_pred_classes, test_data$NEWVITALSTATUS)
print(paste0("Confusion matrix"))
model_1_confusion_matrix
TN = model_1_confusion_matrix[1, 1]
FN = model_1_confusion_matrix[2, 1]
FP = model_1_confusion_matrix[1, 2]
TP = model_1_confusion_matrix[2, 2]
cat("TN:", TN, "FN:", FN, "FP:", FP, "TP:", TP)
Accuracy = (TP + TN)/(TP + FP + TN + FN)
Precision = TP/(TP + FP)
Recall = TP/(TP + FN)
F1_score = 2 * ((Precision * Recall) / (Precision + Recall))
print(paste("Accuracy:", Accuracy))
print(paste("Precision:", round(Precision, 3)))
print(paste("Recall:", round(Recall, 3)))
print(paste("F1_score:", round(F1_score, 3)))
model_1_accuracy = Accuracy
paste("Improvement over naive:")
model_1_accuracy - Naive_Accuracy

```

## Model 2 - STAGE_BEST

```{r}
#create model
model_2 <- glm(NEWVITALSTATUS ~ CANCER_TYPE + STAGE_BEST, data = training_data, family=binomial)
summary(model_2)

#use model to predict outcomes in training data
model_2_train_prob <- model_2 %>%
  predict(type = "response")

#Use model to predict outcomes of test data
model_2_test_prob <- model_2 %>% 
  predict(test_data, type = "response")
model_2_pred_classes <- ifelse(model_2_test_prob > 0.5, 
                                                          "D", 
                                                          "A")

#create confusion matrix
model_2_confusion_matrix <- table(model_2_pred_classes, test_data$NEWVITALSTATUS)
print(paste0("Confusion matrix"))
model_2_confusion_matrix
TN = model_2_confusion_matrix[1, 1]
FN = model_2_confusion_matrix[2, 1]
FP = model_2_confusion_matrix[1, 2]
TP = model_2_confusion_matrix[2, 2]
paste("TN:", TN, "FN:", FN, "FP:", FP, "TP:", TP)
Accuracy = (TP + TN)/(TP + FP + TN + FN)
Precision = TP/(TP + FP)
Recall = TP/(TP + FN)
F1_score = 2 * ((Precision * Recall) / (Precision + Recall))
print(paste("Accuracy:", Accuracy))
print(paste("Precision:", round(Precision, 3)))
print(paste("Recall:", round(Recall, 3)))
print(paste("F1_score:", round(F1_score, 3)))
model_2_accuracy = Accuracy
confint(model_2)
exp(confint(model_2))
```

```{r}
#ggplot(coeff, aes(x = term, y = estimate, fill = term)) + geom_col() + coord_flip()
summary(model_2)$coeff
exp(coefficients(model_2))
```

## Model 3 - T/M/N_BEST

```{r}
#create model
model_3 <- glm(NEWVITALSTATUS ~ CANCER_TYPE + STAGE_BEST + T_BEST + M_BEST + N_BEST, data = training_data, family=binomial)
summary(model_3)

#use model to predict outcomes in training data
model_3_train_prob <- model_3 %>%
  predict(type = "response")

#Use model to predict outcomes of test data
model_3_test_prob <- model_3 %>% 
  predict(test_data, type = "response")
model_3_pred_classes <- ifelse(model_3_test_prob > 0.5, 
                                                          "D", 
                                                          "A")

#create confusion matrix
model_3_confusion_matrix <- table(model_3_pred_classes, test_data$NEWVITALSTATUS)
print(paste0("Confusion matrix"))
model_3_confusion_matrix
TN = model_3_confusion_matrix[1, 1]
FN = model_3_confusion_matrix[2, 1]
FP = model_3_confusion_matrix[1, 2]
TP = model_3_confusion_matrix[2, 2]
paste("TN:", TN, "FN:", FN, "FP:", FP, "TP:", TP)
Accuracy = (TP + TN)/(TP + FP + TN + FN)
Precision = TP/(TP + FP)
Recall = TP/(TP + FN)
F1_score = 2 * ((Precision * Recall) / (Precision + Recall))
print(paste("Accuracy:", Accuracy))
print(paste("Precision:", round(Precision, 3)))
print(paste("Recall:", round(Recall, 3)))
print(paste("F1_score:", round(F1_score, 3)))
model_3_accuracy = Accuracy
```

## Model 4 - Sex

```{r}
#create model
model_4 <- glm(NEWVITALSTATUS ~ CANCER_TYPE + STAGE_BEST + T_BEST + M_BEST + N_BEST+ SEX.x, data = training_data, family=binomial)
summary(model_4)

#use model to predict outcomes in training data
model_4_train_prob <- model_4 %>%
  predict(type = "response")

#Use model to predict outcomes of test data
model_4_test_prob <- model_4 %>% 
  predict(test_data, type = "response")
model_4_pred_classes <- ifelse(model_4_test_prob > 0.5, 
                                                          "D", 
                                                          "A")

#create confusion matrix
model_4_confusion_matrix <- table(model_4_pred_classes, test_data$NEWVITALSTATUS)
print(paste0("Confusion matrix"))
model_4_confusion_matrix
TN = model_4_confusion_matrix[1, 1]
FN = model_4_confusion_matrix[2, 1]
FP = model_4_confusion_matrix[1, 2]
TP = model_4_confusion_matrix[2, 2]
paste("TN:", TN, "FN:", FN, "FP:", FP, "TP:", TP)
Accuracy = (TP + TN)/(TP + FP + TN + FN)
Precision = TP/(TP + FP)
Recall = TP/(TP + FN)
F1_score = 2 * ((Precision * Recall) / (Precision + Recall))
print(paste("Accuracy:", Accuracy))
print(paste("Precision:", round(Precision, 3)))
print(paste("Recall:", round(Recall, 3)))
print(paste("F1_score:", round(F1_score, 3)))
model_4_accuracy = Accuracy
```

## Model 5 - Age

```{r}
#create model
model_5 <- glm(NEWVITALSTATUS ~ CANCER_TYPE + STAGE_BEST + T_BEST + M_BEST + N_BEST+ AGE, data = training_data, family=binomial)
summary(model_5)

#use model to predict outcomes in training data
model_5_train_prob <- model_5 %>%
  predict(type = "response")

#Use model to predict outcomes of test data
model_5_test_prob <- model_5 %>% 
  predict(test_data, type = "response")
model_5_pred_classes <- ifelse(model_5_test_prob > 0.5, 
                                                          "D", 
                                                          "A")

#create confusion matrix
model_5_confusion_matrix <- table(model_5_pred_classes, test_data$NEWVITALSTATUS)
print(paste0("Confusion matrix"))
model_5_confusion_matrix
TN = model_5_confusion_matrix[1, 1]
FN = model_5_confusion_matrix[2, 1]
FP = model_5_confusion_matrix[1, 2]
TP = model_5_confusion_matrix[2, 2]
paste("TN:", TN, "FN:", FN, "FP:", FP, "TP:", TP)
Accuracy = (TP + TN)/(TP + FP + TN + FN)
Precision = TP/(TP + FP)
Recall = TP/(TP + FN)
F1_score = 2 * ((Precision * Recall) / (Precision + Recall))
print(paste("Accuracy:", Accuracy))
print(paste("Precision:", round(Precision, 3)))
print(paste("Recall:", round(Recall, 3)))
print(paste("F1_score:", round(F1_score, 3)))
model_5_accuracy = Accuracy
```

## Model 6 - Deprivation Quintile

```{r}
#create model
model_6 <- glm(NEWVITALSTATUS ~ CANCER_TYPE + STAGE_BEST + QUINTILE_2015, data = training_data, family=binomial)
summary(model_6)
exp(coefficients(model_6))

#use model to predict outcomes in training data
model_6_train_prob <- model_6 %>%
  predict(type = "response")

#Use model to predict outcomes of test data
model_6_test_prob <- model_6 %>% 
  predict(test_data, type = "response")
model_6_pred_classes <- ifelse(model_6_test_prob > 0.5, 
                                                          "D", 
                                                          "A")

#create confusion matrix
model_6_confusion_matrix <- table(model_6_pred_classes, test_data$NEWVITALSTATUS)
print(paste0("Confusion matrix"))
model_6_confusion_matrix
TN = model_6_confusion_matrix[1, 1]
FN = model_6_confusion_matrix[2, 1]
FP = model_6_confusion_matrix[1, 2]
TP = model_6_confusion_matrix[2, 2]
paste("TN:", TN, "FN:", FN, "FP:", FP, "TP:", TP)
Accuracy = (TP + TN)/(TP + FP + TN + FN)
Precision = TP/(TP + FP)
Recall = TP/(TP + FN)
F1_score = 2 * ((Precision * Recall) / (Precision + Recall))
print(paste("Accuracy:", Accuracy))
print(paste("Precision:", round(Precision, 3)))
print(paste("Recall:", round(Recall, 3)))
print(paste("F1_score:", round(F1_score, 3)))
model_6_accuracy = Accuracy
#confint(model_6)
#exp(confint(model_6))
```

# Comparing model accuracies

```{r}
model_name <- c("Naive model", "model_1", "model_2", "model_3", "model_4", "model_5")
model_acc <- round(c(Naive_Accuracy, model_1_accuracy, model_2_accuracy, model_3_accuracy, model_4_accuracy, model_5_accuracy), 3) 
new_feature <- c(NA, "Cancer type", "Stage", "T/M/N", "Sex", "Age")
df <- tibble(model_name, model_acc, new_feature)
df <- df %>%
  mutate(difference = round(model_acc - Naive_Accuracy, 2))
df %>% arrange(desc(model_acc))


```
